Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2020-04-15 Nicolas.tex
Change on name, Yannick corr, soon Greg2+Saul+Nordin

\usepackage{color}
\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
\newcommand{\mri}[1]{{\textcolor{red}{#1}}}
\newcommand{\nn}[1]{{\textcolor[rgb]{1, 0.27, 0}{#1}}}
\newcommand{\yc}[1]{{\textcolor{blue}{#1}}}

@@

\author{N. Nicolas
\thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
\inst{1}
    \and M. Rigault
    \thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
    \inst{2}
    \and Y. Copin\inst{1}   
    \and R. Graziani\inst{2}
    %HERE ALPHABETIC ORDER
    \and M. Briday\inst{1}
    \and Y.-L. Kim\inst{1}
    \and Saul Perlmutter\inst{3}
    \and G. Smadja\inst{1}   
}

@@

\abstract{
{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
measure of survey statistic increase, the question of astrophysical systematic
uncertainties rises, and notably that of the SN~Ia population redshift drift.}
{In this paper, we study the redshift evolution of SN~Ia
\textsc{\texttt{SALT2.4}} lightcurve stretches, which is a purely intrinsic SN
property, to probe this drift.}
{The SN stretch has been shown to strongly correlate with the SN's environment,
and notably with stellar age tracers. Following the prediction of the fraction
of young and old SNe~Ia in nature from \cite{rigault2018}, and assuming
non-drifting underlying stretch distribution for each, we model the expected
SN~Ia stretch distribution as a function of redshift. We test our prediction
against data from the literature chosen to have negligible selection
effects to ensure that any observed drift is indeed astrophysical and not
artificial.}
{We clearly demonstrate that the underlying SNe~Ia stretch distribution is
evolving as a function of redshift, and that the young/old drifting model is a
much better description of the data than any non-drifting model, including the
sample-based asymmetric distributions used by the Beams with Bias Correction
algorithm, used to correct Malmquist bias. Our favored underlying stretch
model is bimodal: a high-stretch mode shared by both young and old environment,
and a low-stretch mode only existing in old environments.}
{The underlying SNe~Ia population is evolving as a function of redshift and the
exact impact on cosmology remains to be studied. Yet, the astrophysical drift of
the SN stretch distribution does affect current Malmquist bias corrections and
thereby distances derived from SN affected by selection effects. We highlight
that such a bias will increase with surveys covering larger redshift ranges,
which is particularly important for LSST.}

@@

\section{Introduction}
Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate $z<0.5$, when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNe~Ia are anchored on Cepheids \citep[the SH0ES
project][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
by cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM, or when the SN luminosity is anchored at intermediate
redshift by the baryon acoustic oscillation (BAO) scale
\citep{riess2019,reid2019,planck2016, feeney2019}. While using the tip of the
red giant branch technique in place of the Cepheids seem to favor an
intermediate value of H$_0$ \citep{freedman2019,freedman2020}, time delay
measurements from strong lensing seem to also favor high H$_0$ values
\citep{wong2019}.

The H$_0$ tension has received a lot of attention as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
tension when accounting for all other probes \citep{knox2019} and the current
most promising scenario appears to be a burst of expansion at the
matter-radiation decoupling moment caused by a (fine tuned) early dark energy
\citep{poulin2019}.

Alternatively, systematic effects affecting one or several of the aforementioned
analysis could also explain the tension.  In \cite{rigault2015} we suggested
that SNe Ia from the Cepheid calibrator sample differs by construction from  the
Hubble flow as it strongly favors young stellar population environments where
one could find Cepheids. This selection effect would impact the derivation of
H$_0$ if SNe~Ia from young and older environments differ in average magnitudes.

For the last decades, numerous analysis have studied the relation between SNe~Ia
and host properties, finding first that the standardized SNe~Ia magnitude
significantly depends on the host stellar-mass, SNe~Ia from high-mass host being
brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
betoule2014, rigault2018, kim19}.  This mass-step correction is currently used
in cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including
for deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
between the SN and the host remains unclear when using global properties such as
the host stellar mass, which raises the question of the accuracy of the
correction. More recently, studies have used the local SN environment to probe
more direct connections between the SN and its environments \citep{rigault2013},
showing that local age-tracers such as the local specific star formation rate or
the local color are more strongly correlated with the standardized SN magnitude
\citep{rigault2018, roman2018, kim18}, suggesting age as the driving parameter
underlying the mass-step. If true, this would have a significant impact for
cosmology because the redshift drift magnitude correction to apply might
strongly vary \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
importance of local SN environmental study remains highly debated
\cite[e.g.][]{jones2015,jones2019} and especially the impact it has on the
derivation of H$_0$ \citep{jones2015,riess2016,riess2018,rose2019}.

In this paper, we take a step aside to probe the validity of our modeling of the
SN population which we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2018}: one old and one younger, the former
having on average lower lightcurve stretches and being brighter after
standardization. We use the correlation between the SN age probed with local
specific star formation rate in \cite{rigault2018} and the SN stretch to model
the expected evolution of the underlying SN stretch distribution as a function
of redshift. This modeling relies on three assumptions: (1) there are two
populations of SNe~Ia; (2) the relative fraction of each of these populations as
a function of redshift follows the age-modeling presented in \cite{rigault2018}
and (3) the underlying distribution of stretch for each age-sample can be
modeled and does not significantly drift. This paper aims at testing this with
datasets from the literature.

The concept of the SNe~Ia age dichotomy arose with study of the the SN~Ia rate.
\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
the relative SNe~Ia rate in galaxies could only be explained if two populations
existed, one young, following the host star formation activity, and one old
following the host stellar mass (the so called “prompt and delayed“ or “A+B“
model). In \cite{rigault2018} we used the specific Star Formation Rate at the SN
location (Local sSFR or LsSFR) to classify which are the younger (those with a
high LsSFR) and which are the older (those with low LsSFR). Furthermore, since
the first SNe~Ia host analyses, the SN stretch has been known to be strongly
correlated with the SN host properties \citep{hamuy1996, hamuy2000}, which has
been extensively confirmed since \citep[e.g.][]{neill2009, sullivan2010,
    lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013, rigault2013,
pan2014, kim19}. Following the “A+B“ model and the connection between SN stretch
and host properties, \cite{howell2007} first discussed the potential redshift
drift of the SN stretch distribution. In this paper we revisit this analysis
with the most recent SNe~Ia dataset.

We present in section~\ref{sec:sample} the sample we are using for this
analysis, which is based on the Pantheon dataset \citep{scolnic2018a}. We
discuss the importance of obtaining a ``complete'' sample, i.e. representative of
the true underlying SNe Ia distribution, and how we build one from the Pantheon
sample.  We then present in section~\ref{sec:modeling} our modeling of the
distribution of stretch as a function of redshift based models for the stretch
distribution of young and old SNe~Ia. Our results are presented in
section~\ref{sec:results}, where we test whether the SN stretch does evolve as a
function of redshift and if the age model is in good agreement with this
evolution. Consequences for cosmology of our results are briefly discussed in
section~\ref{sec:bbc} and we conclude in section~\ref{sec:ccl}.

@@

We base our analysis on the latest SNe~Ia compilation: the Pantheon catalog from
\cite{scolnic2018a}. A naive approach to test the SN stretch redshift drift
would be to simply compare the observed SN stretch distributions in a few bins
of redshift.  In practice, selection effects are affecting the observed SN
stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
with the lightcurve stretch (and color), the first SNe~Ia that a
magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
Consequently, if selection effects are not accounted for, one might confuse true
population drift with selection effects, and conversely.

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring typing
and host redshift, the selection effects of magnitude-limited surveys should be
negligible below a given redshift where even the faintest SNe~Ia can be observed.
In contrast, targeted surveys have highly complex selection functions and will
be discarded from our analysis. Fortunately, modern SN cosmology samples such as
the Pantheon dataset are now dominated by magnitude limited surveys.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
        lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples from
        the Pantheon catalog.  The individual SNe are shown as blue dots and a
        2D histogram is shown in gray to highlight point density. The ellipse
        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the corresponding
        standardized absolute magnitude using the $\alpha$ and $\beta$
        coefficients from \cite{scolnic2018a}. The grey diagonal lines represent
    the $(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and
$z=1.70$ using SNLS's $m_{lim}$ of $24.8$~mag.}
    \label{fig:maglim}
\end{figure}

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color
distributions of SNe~Ia from the following surveys: PanStarrs
\citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
\citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full distribution
\citep{guy2007,betoule2014}; see also \citet{bazin2011} and \citet{campbell2013}
for similar contours, the second using a more conservative $|c| \leq 0.2$ cut.
Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is $M_0=-19.36$
\citep{kessler2009,scolnic2014}, we can derive the absolute standardized
magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$ along the
aforementioned ellipse given the standardization coefficient $\alpha=0.156$ and
$\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is that with
$(x_1=-1.66, c=0.25)$ and an absolute standardized magnitude at peak in Bessel-B
band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect this object
typically a week before and 10 days after peak to build a suitable lightcurve,
the limiting standardized absolute magnitude effectively is approximately
$M_{lim} = -18.00$~mag. Hence, given the magnitude limit $m_{lim}$ of a
magnitude limited survey, one can derive the maximum redshift $z_{lim}$ above
which the faintest SNe~Ia will be missed using the relation between apparent
magnitude, redshift and absolute magnitude $\mu(z_{lim}) = m_{lim} - M_{lim}$.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia  from
        the SDSS, PS1 and SNLS dataset respectively (data from  Pantheon
        \citealt{scolnic2018a}).  The colored parts represent the distribution
        of SNe~Ia kept in our analysis for they are supposedly free from
        selection bias (see section~\ref{sec:sample}). The dark (resp. light)
    color responds to the conservative (resp. fiducial) selection cut.}
    \label{fig:cuts}
\end{figure}

SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
filter that has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.  Fig.~14 of
\citealt{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both will be
considered, as discussed below.

Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure could also
suggest a more conservative $z_{lim}$ of 0.27.

In similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
SDSS surveys were more sensitive to limited spectroscopic resources.
\cite{kessler2009} pointed out that during year-1 of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to a
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were used, and \cite{kessler2009} and \cite{dilday2008}
show a relative completeness up to $z_{lim}=0.2$. Following these analyses we
will use $z_{lim}=0.2$ as the baseline SDSS redshift limit.

For the rest of the analysis, we will consider the aforementioned redshift
limits as well as more conservative cuts for each of these surveys, namely
$z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS1 and $z_{lim}=0.55$ for SNLS.

The complete sample selection is summarized in Table~\ref{tab:sample}, and the
redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
As expected, the selected redshift limits roughly correspond to the peak of
these histograms.

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
4.2.2). Their spectroscopic follow-up was done over a redshift range of
$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
deeper. As such, these SNe~Ia should also be a random sampling of the underlying
SN population. The SNfactory sample is particularly useful for studying SN property
drift as it enables us to have a large SN~Ia sample at $z<0.1$.

Finally, we included the HST sample from Pantheon, that follows the same logic
of having a search deeper than the follow-up and we therefore kept it entirely
\citep{strolger04}, see Table~\ref{tab:sample}. We present the stretch
distribution and redshift histogram of these five surveys in
Fig.~\ref{fig:sample}.

\begin{table}
    \centering
    \caption{Composition of the SNe~Ia dataset used in this analysis.
    Conservative cuts are indicated in brackets.}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & z$_{lim}$ & N$_{\mathrm{SN}}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & -- & 114\\[0.30em]
        SDSS & 0.20(0.15) & 167(82)\\[0.30em]
        PS1 & 0.30(0.27) & 160(122)\\[0.30em]        
        SNLS & 0.60(0.55) & 102(78)\\[0.30em]
        HST & -- & 26\\[0.30em]
        \hline\\[-0.8em]
        Total & -- & 569(422)\\[0.30em]
        \hline
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb-cividis.pdf}
    \caption{\textit{Lower}: \textsc{\texttt{SALT2.4}} lightcurve stretch as a
        function of redshift (in log) for each survey considered in this
        analysis (see legend).  Full (resp. empty) markers correspond to the
        conservative (resp. fiducial) selection cuts. \textit{Top}: combined
        redshift histogram in dark (resp. light) gray for the conservative
    (resp. fiducial) selection cuts.}
    \label{fig:sample}
\end{figure}

>>>>

\usepackage[dvipsnames]{xcolor}
\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
\newcommand{\mri}[1]{{\textcolor{red}{#1}}}
\newcommand{\nn}[1]{{\textcolor[rgb]{0.25, 0.50, 0}{#1}}}
\newcommand{\yc}[1]{{\textcolor{BrickRed}{#1}}}

@@

\author{N. Nicolas
\thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
\inst{1}
    \and M. Rigault
    \thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
    \inst{2}\and\\
    Y. Copin\inst{1}   
    \and R. Graziani\inst{2}
    %HERE ALPHABETIC ORDER
    \and G. Aldering\inst{3}
    \and M. Briday\inst{1}
    \and Y.-L. Kim\inst{1}
    \and Saul Perlmutter\inst{3}
    \and G. Smadja\inst{1}   
}

@@

\abstract{
{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
measure of survey statistic increase, the question of astrophysical systematic
uncertainties rises, and notably that of the SN~Ia population redshift drift.}
{In this paper, we study the redshift evolution of SN~Ia
\textsc{\texttt{SALT2.4}} lightcurve stretch\yc{\sout{es}}, \yc{\sout{which is}} a purely intrinsic SN
property, to probe this drift.}
{The SN stretch has been shown to strongly correlate with the SN's environment,
and notably with stellar age tracers. Following the \yc{evolution of} the fraction
of young and old SNe~Ia \yc{\sout{in nature} as predicted by} \cite{rigault2018}, and assuming
non-drifting \yc{$\to$constant/stationary?}  underlying stretch distribution for each \yc{population}, we model the expected
SN~Ia stretch distribution as a function of redshift. We test our prediction
against data from the literature chosen to have negligible selection
effects to ensure that any observed drift is indeed astrophysical and not
\yc{\sout{artificial} observational}.}
{We clearly demonstrate that the underlying SN\yc{e?}~Ia stretch distribution is
evolving as a function of redshift, and that the young/old drifting model is a
much better description of the data than any non-drifting model, including the
sample-based asymmetric distributions used by the Beams with Bias Correction
algorithm used to correct Malmquist bias. Our favored underlying stretch
model is bimodal: a high-stretch mode shared by both young and old environments,
and a low-stretch mode \yc{\sout{only existing in} exclusive to} old environments.}
{\sout{The underlying SNe~Ia population is evolving as a function of redshift \yc{This is purely a repetition of a previous sentence} \nn{\textit{see new sentence}} and the
exact impact on cosmology remains to be studied.} \nn{The precise impact of the redshift evolution of the underlying SNe Ia population on cosmology remains to be studied}. Yet, the astrophysical drift of
the SN stretch distribution does affect current Malmquist bias corrections and
thereby distances derived from SN affected by selection effects. We highlight
that such a bias will increase with surveys covering larger redshift ranges,
which is particularly important for LSST.}

@@

\section{Introduction}

\yc{YC: you should run a grammar corrector, e.g. \texttt{grammarly}. e.g. "SNe dichotomy" vs. "SN dichotomy", "analysis" vs. "analyses", etc.}

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate ($z<0.5$), when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNe~Ia are anchored on Cepheids \citep[the SH0ES
project,][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
\yc{from} cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM \yc{ref?} \nn{textit{isn't it at the end of the sentence? Do you think it should be before the "or"?}}, or when the SN luminosity is anchored at intermediate
redshift by the baryon acoustic oscillation (BAO) scale
\citep{riess2019,reid2019,planck2018, feeney2019}. While using the tip of the
red giant branch technique in place of the Cepheids seem to favor an
intermediate value of H$_0$ \citep{freedman2019,freedman2020}, time delay
measurements from strong lensing seem to also favor high H$_0$ values
\citep{wong2019}.

The H$_0$ tension has received a lot of attention as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
tension when accounting for all other probes \citep{knox2019} and the current
most promising scenario appears to be a burst of expansion at the
matter-radiation decoupling moment caused by a (fine tuned) early dark energy
\citep{poulin2019}. \yc{I don't see the point of going into this debate...}

Alternatively, systematic effects affecting one or several of the aforementioned
analyses could also explain the tension. \cite{rigault2015} suggested
that SNe Ia from the Cepheid calibrator sample differ by construction from the
Hubble flow \yc{ones} as it \yc{they?} strongly favors \yc{SNe don't favor anything...} \nn{\textit{but the Cepheid sample does, right?}} young stellar population environments where
one could find Cepheids. This selection effect would impact the derivation of
H$_0$ if SNe~Ia from young and older environments differ in average \yc{standardized} magnitudes.

For the last decades, numerous analyses have studied the relation between SNe~Ia
and host properties, finding first that the standardized SNe~Ia magnitude
significantly depends on the host stellar mass, SNe~Ia from high-mass host being
brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
betoule2014, rigault2018, kim19}.  This mass-step correction is currently used
in cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including
for deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
between the SNe and their host remains unclear when using global properties such as
the host stellar mass, which raises the question of the accuracy of the
correction. More recently, studies have used the local SN environment to probe
more direct connections between the SN and its environments \citep{rigault2013},
showing that local age tracers such as the local specific star formation rate or
the local color are more strongly correlated with the standardized SN magnitude
\citep{rigault2018, roman2018, kim18}, suggesting age as the driving parameter
underlying the mass-step. If true, this would have a significant impact for
cosmology because the redshift drift magnitude correction to apply might
strongly vary \yc{just the existence of a redshift-drift mag. correction is of importance here, even if it does not vary} \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
importance of local SN environmental studies remains highly debated
\cite[e.g.][]{jones2015,jones2019} and especially the impact it \yc{what does `it' refer to, the `importance'?} \nn{\textit{no, to the SN environment. Do you think it needs rephrasing?}} has on the
derivation of H$_0$ \citep{jones2015,riess2016,riess2018,rose2019}.

In this paper, we take a step aside to probe the validity of our modeling of the
SN population which we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2018}: one old and one younger, the former
having on average lower lightcurve stretches and being brighter after
standardization. We use the correlation between the SN age, \yc{as} probed by local
specific star formation rate \citep{rigault2018}, and the SN stretch to model
the expected evolution of the underlying SN stretch distribution as a function
of redshift. This modeling relies on three assumptions: (1) there are two
\yc{distinct} populations of SNe~Ia; (2) the relative fraction of each of these populations as
a function of redshift follows the \yc{model} presented in \cite{rigault2018}
and (3) the underlying distribution of stretch for each age sample \yc{is constant} \nn{\textit{what does it mean for a distribution to be constant? It seems weird that we say "does not significantly evolve" after saying it is constant, no?}} and does not significantly \yc{evolve with time}. This paper aims at testing this \yc{specific model} with
datasets from the literature.

\yc{Shouldn't this paragraph, introducing the concept of old and young populations and lsSFR, be moved \emph{before} the previous one?}
The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate.
\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
the relative SNe~Ia rate in galaxies could be explained if two populations
existed, one young, following the host star formation activity, and one old
following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
model). In \cite{rigault2018} we used the specific Star Formation Rate at the SN
location (Local sSFR or LsSFR \yc{this has already been used multiple times, define before}) to classify which are the younger (those with a
high LsSFR) and which are the older (those with low LsSFR). Furthermore, since
the first SNe~Ia host analyses, the SN stretch has been known to be strongly
correlated with the SN host properties \citep{hamuy1996, hamuy2000}, which \yc{what does this ``which'' refer to?} has
been extensively confirmed since \citep[e.g.][]{neill2009, sullivan2010,
    lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013, rigault2013,
pan2014, kim19}. Following the ``A+B'' model and the connection between SN stretch
and host properties, \cite{howell2007} first discussed the potential redshift
drift of the SN stretch distribution. In this paper we revisit this analysis
with the most recent SNe~Ia dataset.

We present in section~\ref{sec:sample} the sample we are using for this
analysis, \yc{derived from} the Pantheon catalog \citep{scolnic2018a}. We
discuss the importance of obtaining a ``complete'' sample, i.e. representative of
the true underlying SNe Ia distribution, and how we build one from the Pantheon
sample. We then present in section~\ref{sec:modeling} our modeling of the
distribution of stretch as a function of redshift\nn{-}based models for the stretch
distribution of young and old SNe~Ia \yc{what??? rephrase} \nn{\textit{what isn't clear? Does the hyphen help?}}. Our results are presented in
section~\ref{sec:results}, where we test whether the SN stretch \yc{distribution} evolves as a
function of redshift and if the \yc{aforementioned} age model is in good agreement with this
evolution. Consequences for cosmology of our results are briefly discussed in
section~\ref{sec:bbc} and we conclude in section~\ref{sec:ccl}.

@@

We base our analysis on the \yc{most recent comprehensive} SNe~Ia compilation, the Pantheon catalog from
\cite{scolnic2018a}. A naive approach to test the SN stretch redshift drift
would be to simply compare the observed SN stretch distributions in a few bins
of redshift.  In practice, \yc{however}, \yc{differential} selection effects are affecting the observed SN
stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
with the lightcurve stretch (and color), the first SNe~Ia that a
magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
Consequently, if \yc{magnitude-related} selection effects are not accounted for, one might confuse true
population drift with \yc{\sout{selection effects} survey properties}, and conversely.

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring typing
and host redshift, the selection effects of magnitude-limited surveys should be
negligible below a given redshift \yc{at which} even the faintest SNe~Ia can be observed.
In contrast, targeted surveys have highly complex selection functions and will
be discarded from our analysis. Fortunately, modern SN cosmology samples such as
the Pantheon \yc{one} are now dominated by magnitude-limited surveys.

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia from the following surveys: PanStarrs
\citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
\citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full distribution
\citep{guy2007,betoule2014}; see also \citet{bazin2011} and \citet{campbell2013}
for similar contours, the second using a more conservative $|c| \leq 0.2$ cut.
Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is $M_0=-19.36$
\citep{kessler2009,scolnic2014}, we can derive the absolute standardized
magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$ along the
aforementioned ellipse given the standardization coefficient $\alpha=0.156$ and
$\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is that with
$(x_1=-1.66, c=0.25)$ and an absolute standardized magnitude at peak in Bessel $B$
band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect this object
typically a week before and 10 days after peak to build a suitable lightcurve,
the \yc{effective} limiting standardized absolute magnitude is approximately
$M_{lim} = -18.00$~mag. Hence, given the magnitude limit $m_{lim}$ of a
magnitude limited survey, one can derive the maximum redshift $z_{lim}$ above
which the faintest SNe~Ia will be missed using the relation between apparent
magnitude, redshift and absolute magnitude $\mu(z_{lim}) = m_{lim} - M_{lim}$.

% YC: Figs and Tables *after* their 1st ref. in the text!
\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
        lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples of
        the Pantheon catalog.  The individual SNe are shown as blue dots and a
        2D histogram is shown in gray to highlight point density \yc{I don't think it's really needed}. The ellipse
        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the corresponding
        standardized absolute magnitude using the $\alpha$ and $\beta$
        coefficients from \cite{scolnic2018a}. The grey diagonal lines represent
    the $(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and
$z=1.70$ using SNLS's $m_{lim}$ of $24.8$~mag.}
    \label{fig:maglim}
\end{figure}

SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, that has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.  Fig.~14 of
\cite{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both \yc{limits} will be
considered, as discussed below.

Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure \yc{is also suggestive of} a more conservative $z_{lim}=0.27$.

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
SDSS surveys were more sensitive to limited spectroscopic resources;
\cite{kessler2009} pointed out that during \yc{first} year of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow-up, corresponding to a
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were \yc{available}, and \cite{kessler2009} and \cite{dilday2008}
show a relative completeness \yc{what is a \emph{relative completeness}???} up to $z_{lim}=0.2$. Following these analyses, we
will use $z_{lim}=0.2$ as the baseline SDSS redshift limit.

For the rest of the analysis, we will consider the aforementioned redshift
limits, as well as more conservative \yc{values} for each of these surveys, namely
$z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS1 and $z_{lim}=0.55$ for SNLS. \yc{why do you remind the conservative values, and not the fiducial ones? Just refer to Table~1.}

The complete sample selection is summarized in Table~\ref{tab:sample}, and the
redshift distribution of these three surveys are \yc{is?} shown in Fig.~\ref{fig:cuts}.
As expected, the selected redshift limits roughly correspond to the peak of
these histograms.

\begin{table}
    \centering
    \caption{Composition of the SNe~Ia dataset used in this analysis.
    Conservative cuts are indicated in brackets.}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline%\\[-0.8em] % YC: don't, it's not your business to do that!
        Survey & $z_{lim}$ & $N_{\mathrm{SN}}$ \\%[0.15em]
        \hline%\\[-0.8em]
        SNf & -- & 114\\%[0.30em]
        SDSS & 0.20 (0.15) & 167 (82)\\%[0.30em]
        PS1 & 0.30 (0.27) & 160 (122)\\%[0.30em]        
        SNLS & 0.60 (0.55) & 102 (78)\\%[0.30em]
        HST & -- & 26\\%[0.30em]
        \hline%\\[-0.8em]
        Total & -- & 569 (422)\\%[0.30em]
        \hline
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia from
        the SDSS, PS1 and SNLS dataset respectively (data from Pantheon,
        \citealt{scolnic2018a}).  The colored parts represent the distribution
        of SNe~Ia kept in our analysis for they are supposedly free from
        selection bias (see section~\ref{sec:sample}). The darker (resp. lighter)
    color responds to the conservative (resp. fiducial) selection cut.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3 and~4.2.2; \mr{SNe~Ia time series are published in \citealt{Saunders2020}}). 
The spectroscopic follow-up was done over a redshift range of
$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the search was much
deeper. As such, these SNe~Ia \yc{\sout{should also} are assumed to} be a random sampling of the underlying
SN population. The SNfactory sample is particularly useful for studying SN property
drift, as it enables us to have a large SN~Ia sample at $z<0.1$.

Finally, we include the HST sample from Pantheon, that \yc{\sout{follows the same logic
of having} similarly have} a search deeper than the follow-up and \yc{that} we therefore kept entirely
\citep{strolger04}. %, see Table~\ref{tab:sample}.

We present the stretch
distribution and redshift histogram of these five surveys \yc{up to their respective $z_{lim}$} in
Fig.~\ref{fig:sample}.
